---
title: Stroke Clinical Predictive Model
author: "Tosin Dairo"
date: MAY 25, 2019
output: rmarkdown::github_document
categories: ["R"]
tags: ["Health", "Predictive_Models", "Machine_Learning"]
---

## Clinical Predictive Model classifier for stroke

Stroke-CPM is a tool that takes information available about a patient and their observed factors, and makes a prediction regarding their diagnosis and causal factors.

+ Diagnosis – detects presence of stroke currently
+ Causal Factors – assess whether observed factor(s) is/are likely to be contributors to the development of stroke presently or in the future

<br></br>

#### Observed Factors:

+ Gender
+ Age
+ Hypertension
+ Heart Disease
+ Ever Married
+ Work Type
+ Residence Type
+ Average Glucose Level
+ BMI
+ Smoking Status

<br></br> 

```{r include=FALSE}
getwd()
library(reticulate)
library(mice)
library(tidyverse)
library(ggplot2)
library(MASS)
library(stargazer)
library(sjPlot)
library(dataPreparation)
library(DMwR)
library(randomForest)  
library(caret)
library(pROC)
library(ROCR)
use_python("/Volumes/Loopdisk/Dev/PyDsc/env/bin/python3", required = TRUE)
# use_python("/usr/bin/python3", required = TRUE)
# use_virtualenv("~/.virtualenvs/r-reticulate", required = TRUE)
d <- read.csv("/Volumes/Loopdisk/Stroke-CPM/data/train_2v.csv")
reticulate::py_config()
```



```{python include=FALSE}
import matplotlib
matplotlib.use('TkAgg')
import pandas as pd
import matplotlib.pyplot as plt
import seaborn as sns
```



```{r}
head(d)
```



```{r}
d$smoking_status <- na_if(d$smoking_status, "")
```



```{python}
print(r.d.isnull().sum())
fig, ax = plt.subplots(figsize=(14,12))
sns.heatmap(r.d.isnull(),yticklabels=False,cbar=False,cmap='viridis') 
```



```{r}
d$gender <- as.factor(d$gender)
d$ever_married <- as.factor(d$ever_married)
d$work_type <- as.factor(d$work_type)
d$Residence_type <- as.factor(d$Residence_type)
d$smoking_status <- as.factor(d$smoking_status)
str(d)
```



```{r}
imp <- mice(d, seed = 3333)
```



```{r}
imp$method
```



```{r}
dx <- mice::complete(imp)
```



```{python}
print(r.dx.isnull().sum())
fig, ax = plt.subplots(figsize=(14,12))
sns.heatmap(r.dx.isnull(),yticklabels=False,cbar=False,cmap='viridis') 
```



```{python}
fig, ax = plt.subplots(figsize=(14,12))
sns.pairplot(r.dx)
plt.show()
```



```{python}
sns.countplot(x='stroke', data = r.dx)
```



```{r echo=FALSE}
{{knitr::spin_child('label_encode.R')}}
```




```{r}
result <- encode.fit_transform(dx)

dx_ <- result[[2]]
# encode.transform()
```



```{r}
tab_df(table(dx_$stroke), title = "Stroke Class")
tab_df(table(dx_$hypertension), title = "Hypertension Class")
tab_df(table(dx_$heart_disease), title = "Heart Disease Class")
```



```{r}
# encoding <- build_encoding(dataSet = X_train, cols = "auto", verbose = TRUE)
# X_train <- one_hot_encoder(dataSet = X_train, encoding = encoding, drop = TRUE, verbose = TRUE)
# X_test <- one_hot_encoder(dataSet = X_test, encoding = encoding, drop = TRUE, verbose = TRUE)
```



```{r}
dx_$hypertension <- as.factor(dx_$hypertension)
dx_$heart_disease <- as.factor(dx_$heart_disease)
dx_$stroke <- as.factor(dx_$stroke)
dx_$gender <- as.factor(dx_$gender)
dx_$ever_married <- as.factor(dx_$ever_married)
dx_$work_type <- as.factor(dx_$work_type)
dx_$Residence_type <- as.factor(dx_$Residence_type)
dx_$smoking_status <- as.factor(dx_$smoking_status)
str(dx_)
```



```{r}
## Smote : Synthetic Minority Oversampling Technique To Handle Class Imbalancy In Binary Classification
# balanced.data <- SMOTE(Class ~., dresstrain, perc.over = 4800, k = 5, perc.under = 1000)
set.seed(3333)

balanced_dx_ <-SMOTE(stroke ~., dx_, perc.over = 2100 , k = 5, perc.under = 160)

table(balanced_dx_$stroke)


```



```{r}
# Random sample indexes
train_index <- sample(1:nrow(balanced_dx_), 0.8 * nrow(balanced_dx_))
test_index <- setdiff(1:nrow(balanced_dx_), train_index)

# Build X_train, y_train, X_test, y_test
X_train <- balanced_dx_[train_index, 2:12]
# y_train <- dx_[train_index, "stroke"]

X_test <- balanced_dx_[test_index, 2:12]
# y_test <- dx_[test_index, "stroke"]

table(X_train$stroke)
```



```{r}
# balanced_dx_$hypertension <- as.factor(balanced_dx_$hypertension)
# balanced_dx_$heart_disease <- as.factor(balanced_dx_$heart_disease)
# balanced_dx_$stroke <- as.factor(balanced_dx_$stroke)
# balanced_dx_$gender <- as.factor(balanced_dx_$gender)
# str(balanced_dx_)
set.seed(3333)
model <- glm (stroke ~ ., data=X_train, family = binomial)
summary(model)
```
  


```{r}
stepAIC(model, direction='both')

model_select <- glm(formula = stroke ~ gender + age + hypertension + heart_disease + 
    ever_married + work_type + avg_glucose_level + smoking_status, 
    family = binomial, data = X_train)
summary(model_select)
```



```{r}
## Predict the Values
predict <- predict(model, X_test, type = 'response')

## Create Confusion Matrix
table(X_test$stroke, predict > 0.5)
```



```{r}
#ROCR Curve
ROCRpred <- prediction(predict, X_test$stroke)
ROCRperf <- performance(ROCRpred, 'tpr','fpr')
plot(ROCRperf, colorize = TRUE, text.adj = c(-0.2,1.7))
```



```{r}
auc(roc(X_test$stroke ~ predict))
```



```{r}
set.seed(3333)
rf = randomForest(stroke ~ .,  
                   ntree = 100,
                   data = X_train)
plot(rf) 
```



```{r}
varImp(rf)
```



```{r}
## Important variables according to the model
varImpPlot(rf,  
           sort = T,
           n.var=25,
           main="Variable Importance")
```



```{r}
predicted.response <- predict(rf, X_test)


confusionMatrix(data=predicted.response,  
                reference=X_test$stroke)
```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```



```{r}

```